---
title: "Predictors of Mortality in CLD with renal dysfunction"
author: "Anupam kumar Singh, MD"
date: "2 November 2017"
output: html_document
---

we will try to evaluate factors affecting mortality in HRS .
We have already calculated descritive statistics, done ANOVA of NGAL and done comparative ROC Aanalysis of biomarkers like FeNa,NGAL,Urine Na,Urine Creatinine  which are accessible along with master chart [here](https://drive.google.com/open?id=0B7gX8b7uHJW2LUhhYnF1NVNoUWc)

Let us do survival analysis now, first dplyr::select limited variables

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r}
library(tidyverse)
master= read.csv("master.csv")
master1 = master %>% dplyr::select(diagnoses,MELD,CTP,serum_sodium,NGAL,feNa,age,outcome_30,outcome_90) %>% mutate(time30 = rep(30,length(outcome_30)),censor30=case_when(outcome_30=="d"~1,TRUE~0),time90 = rep(90,length(outcome_90)),censor90=case_when(outcome_30=="d"~1,TRUE~0)) %>% gather(time_period,days,time30,time90) %>% gather(censor_period,status,censor30,censor90)
```


Lets prepare our data for survival analysis

```{r}
master2 = master %>% dplyr::select(diagnoses,MELD,CTP,serum_sodium,NGAL,feNa,age,outcome_30,outcome_90) %>% mutate(days= case_when(outcome_30=="d"~30,TRUE~90),status=case_when(outcome_30=="d"~1,outcome_90=="d"~1,TRUE~0))
```

```{r}
require("survival")
require('survminer')
ngalfit <- survfit(Surv(days, status) ~ 1, data = master2)
summary(ngalfit)

```

survival plot from same

```{r}
ggsurvplot(ngalfit, conf.int=FALSE,data = master2)
```


classifying CTP n MELD for further use in survival curves

```{r}
master3 = master2 %>% filter(!(diagnoses=="Control")) %>% mutate(MELD_CAT = case_when(MELD<=10 ~ "MELD<=10",MELD<=18 ~ "MELD(10-18)",MELD<=24 ~ "MELD(19-24)",MELD>=25~"MELD>=25"),CTP_CATEGORIES= case_when(CTP<7~"A",CTP<10~"B",TRUE~"C"))
```


Plot survival by diagnosetic categories

```{r}
ngalfit1 <- survfit(Surv(days, status) ~ diagnoses, data = master3)
summary(ngalfit1)
```

```{r}
ggsurvplot(ngalfit1, conf.int=FALSE,data = master2)
```

Plot survival by MELD category

```{r}
ngalfit2 <- survfit(Surv(days, status) ~ MELD_CAT, data = master3)
summary(ngalfit2)
```

```{r}
ggsurvplot(ngalfit2, conf.int=FALSE,data = master3)
```


Plot survival by CTP

```{r}
ngalfit3<- survfit(Surv(days, status) ~ CTP_CATEGORIES, data = master3)
summary(ngalfit3)
```

```{r}
ggsurvplot(ngalfit3, conf.int=FALSE,data = master3)
```

Let us do multivariate analysis of mortality with predictive factor

```{r}
fit4 = glm(as.factor(outcome_30)~age+MELD,family=binomial(),data=master)
summary(fit4)
```

Lets add NGAL  to model with MELD only and see improvement in AIC  has been used often

```{r}
fit5 = glm(as.factor(outcome_30)~age+NGAL+MELD,family=binomial(),data=master)
summary(fit5)
```

so adding NGAL to model leads to decrease in AIC (>1) more than implying its an effective variable


```{r}
fit6 = glm(as.factor(outcome_30)~age+NGAL+MELD+serum_sodium,family=binomial(),data=master)
summary(fit6)
```


Adding serum sodium doesnt lead to fall in AIC, so even adding an extra predictor isnt improving predictive power of our model

```{r}
anova(fit4,fit5)
```

Let's do ROC curve analysis of our model with MELD,NGAL

```{r}
master$predpr <- predict(fit5,type=c("response"))
library(pROC)
roccurve <- roc(outcome_30 ~ predpr,data=master)
plot(roccurve)
```

```{r}
auc(roccurve)
```

thus AUC of combined MELD n NGAL model is 0.818

```{r}
library(rms)

master4 = master %>% dplyr::select(outcome_30,NGAL,MELD,age)
master.dd= datadist(master4)


options(datadist="master.dd")

fit7 = lrm(outcome_30~age+NGAL+MELD,data=master4,x=TRUE,y=TRUE)
fit7
```

Nomogram of model

```{r}
plot(nomogram(fit7,fun=function(x)1/(1+exp(-x)),  # or fun=plogis
    fun.at=c(.001,.01,.05,seq(.1,.9,by=.1),.95,.99,.999),
    funlabel="Risk of Death"))
```


Let's do internal validation bootstrap

```{r}
vl = validate(fit7,B=200)
vl
```

so our bootstrapped model shows considerable optimism implying on external validation our model is going to fare worse

lets's see calibration estimates


```{r}
plot(calibrate(fit7,B=200))
```

we can see while negative predictive probability is good the positive predictive of our model fares a bit worse

```{r}
library(strengejacke)

sjp.glm(fit5)
```

Lets calculate adjusted odds ratio from our final model

```{r}
require(MASS)
exp(cbind(coef(fit5), confint(fit5)))  
```

